04_DESeq_sexes

### 2019-05-24 analysis/04_DESeq2_sexes.Rmd
new update to pca and total DEGs

# load custom functions  
source("../R/functions.R")

@@

a.colData$xlabel <- a.colData$treatment
levels(a.colData$xlabel) <-  c("control", "nest building", "egg lay",
                               "eggs early", "eggs early remove", 
                               "eggs mid", "eggs mid hatch", "eggs mid remove",
                               "eggs end", "eggs end remove",
                               "chicks hatch",  "chicks hatch remove"  ,
                               "chicks mid", "eggs delay", "eggs delay hatch", "chicks end")

a.colData$lastday <- factor(a.colData$lastday, levels =  c("control", "nest building", "eggs lay",
                               "eggs early",  
                               "eggs middle", 
                               "eggs later", 
                               "eggs delay",
                               "chicks hatch",
                               "chicks early", 
                               "chicks later", 
                               "empty nest"))

@@

colorlastday <-  c("control" = "#636363", 
                   "nest building" = "#fb6a4a", 
                   "eggs lay" = "#c7e9c0",
                   "eggs early" = "#a1d99b",
                   "eggs middle" = "#74c476", 
                   "eggs later" = "#31a354",  
                   "eggs delay" = "#006d2c",
                   "chicks hatch" = "#cbc9e2",   
                   "chicks early" = "#9e9ac8", 
                   "chicks later" = "#6a51a3",
                   "empty nest" = "#a50f15")

colorpenultimate <- c("control" = "#636363", 
                      "nest building" = "#fb6a4a", 
                      "eggs early" = "#a1d99b",
                   "eggs middle" = "#74c476", 
                   "eggs later" = "#31a354",  
                   "eggs delay" = "#006d2c",
                   "chicks hatch" = "#cbc9e2",   
                   "chicks early" = "#9e9ac8", 
                   "chicks later" = "#6a51a3")

colorlastdaybad <-  c("control" = "#636363", 
                   "nest bldg" = "#fb6a4a", 
                   "lay" = "#c7e9c0",
                   "incubate begin" = "#a1d99b",
                   "incubate middle" = "#74c476", 
                   "incubate end" = "#31a354",  
                   "eggs late" = "#006d2c",
                   "nestling being" = "#cbc9e2",  
                   "nestling begin" = "#cbc9e2",   
                   "nestling middle" = "#9e9ac8", 
                   "nestling end" = "#6a51a3",
                   "empty nest" = "#a50f15")

colorpenultimatebad <- c("control" = "#636363", 
                      "nestbldg" = "#fb6a4a", 
                      "incubate begin" = "#a1d99b",
                   "incubate middle" = "#74c476", 
                   "incubate end" = "#31a354",  
                   "eggs late" = "#006d2c",
                   "nestling begin" = "#9e9ac8",   
                   "nestling middle" = "#6a51a3", 
                   "nestling end" = "#3f007d")

summary(a.colData[c(7,3,4,5,8,9, 10,11)])

@@

```{r totalDEGs, eval = F}

#create list of groups for deseq contrasts
group1 <- levels(a.colData$treatment)
group2 <- group1

a <- plottotalDEGs(dds.female_hypothalamus, "female hypothalamus")
b <- plottotalDEGs(dds.female_pituitary, "female pituitary")
c <- plottotalDEGs(dds.female_gonad, "female gonad")
d <- plottotalDEGs(dds.male_hypothalamus, "male hypothalamus")
e <- plottotalDEGs(dds.male_pituitary, "male pituitary")
f <- plottotalDEGs(dds.male_gondad, "male gonad")

plot_grid(a + theme(legend.position = "none"),
          b + theme(legend.position = "none"),
          c,
          d + theme(legend.position = "none"),
          e + theme(legend.position = "none"),
          f,
          nrow = 2, rel_widths = c(0.3, 0.3, 0.4)) 
```

@@

```{r pcadf}
p1 <- plotPCAs(dds.female_hypothalamus, "female hypothalamus")
p2 <- plotPCAs(dds.female_pituitary, "female pituitary")      
p3 <- plotPCAs(dds.female_gonad, "female gonad")
p4 <- plotPCAs(dds.male_hypothalamus, "male hypothalamus")
p5 <- plotPCAs(dds.male_pituitary, "male pituitary")
p6 <- plotPCAs(dds.male_gondad, "male gonad")
p1
p2
p3
p4
p5
p6

@@

```{r pca, fig.dim = c(12, 10), eval=T}
mylegend <- get_legend(p1)

>>>>

source("../R/functions.R")  # load custom functions 
source("../R/themes.R")  # load custom themes and color palletes

@@

a.colData$xlabel <- a.colData$treatment

levels(a.colData$xlabel) <-  c("control", "nest.building", "egg.lay",
                               "eggs.early", "eggs.early.remove", 
                               "eggs.mid", "eggs.mid.hatch", "eggs.mid.remove",
                               "eggs.end", "eggs.end.remove",
                               "chicks.hatch",  "chicks.hatch.remove"  ,
                               "chicks.mid", "eggs.delay", "eggs.delay.hatch", "chicks.end")

a.colData$lastday <- factor(a.colData$lastday, levels =  c("control", "nest building", "eggs lay",
                               "eggs early", "eggs middle",  "eggs later","eggs delay",
                               "chicks hatch", "chicks early",  "chicks later", "empty nest"))

@@

summary(a.colData[c(7,3,4,5,8,9, 10,11)])
```

@@

```{r totalDEGs, eval = T}
#create list of groups for deseq contrasts
group1 <- levels(a.colData$treatment)
group2 <- group1

a <- plottotalDEGs(dds.female_hypothalamus, "female hypothalamus")
b <- plottotalDEGs(dds.female_pituitary, "female pituitary")
c <- plottotalDEGs(dds.female_gonad, "female gonad")
d <- plottotalDEGs(dds.male_hypothalamus, "male hypothalamus")
e <- plottotalDEGs(dds.male_pituitary, "male pituitary")
f <- plottotalDEGs(dds.male_gondad, "male gonad")
```

```{r plottotalDEGs, eval = T}
plot_grid(a + theme(legend.position = "none"),
          b + theme(legend.position = "none"),
          c + theme(legend.position = "none"),
          d + theme(legend.position = "none"),
          e + theme(legend.position = "none"),
          f + theme(legend.position = "none"),
          nrow = 2) 
```

@@

```{r pcadf, eval=F}
p1 <- plotPCAs(dds.female_hypothalamus, "female hypothalamus")
p2 <- plotPCAs(dds.female_pituitary, "female pituitary")      
p3 <- plotPCAs(dds.female_gonad, "female gonad")
p4 <- plotPCAs(dds.male_hypothalamus, "male hypothalamus")
p5 <- plotPCAs(dds.male_pituitary, "male pituitary")
p6 <- plotPCAs(dds.male_gondad, "male gonad")
p1
p2
p3
p4
p5
p6

@@

```{r pca, fig.dim = c(12, 10), eval=F}
mylegend <- get_legend(p1)
